---
title: "chapter-3"
output: html_document
date: "2023-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Solutions to some of the exercises in Chapter 3

<br>

3.8: Suppose $\gamma$ = 0.5, and the following sequence of rewards is received $R_{1}$ = -1, $R_{2}$ = 2, $R_{3}$ = 6, $R_{4}$ = 3, & $R_{5}$ = 2, with T=5. What are $G_{0}$, $G_{1}$, ..., $G_{5}$? 
<br>

```{r}
T = 5
Gs <- rep(0,T+1) # T+1 because we're also estimating G0
Rs <- c(-1, 2, 6, 3, 2)
gamma = 0.5

# Gs[5] = 0 because G_{T} = 0
for (i in seq(T, 1, -1)){ # go from 4 to 0
Gs[i] = Rs[i] + (gamma*Gs[i+1]) 
}

Gs

```
<br>

Take home is that all of the future is packaged into the estimate for the next time-step.

<br>

3.9 Suppose $\gamma$ = 0.9, and the reward sequence is $R_{1}$ = 2, followed by an infinite sequence of 7s. What are $G_{1}$ and $G_{0}$?

Here we can use formula 3.8:

$G_{t} = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+3} + ... = \displaystyle\sum_{k=0} ^{\infty} \gamma^k R_{t+k+1}$

Instead we have:

$G_{1} = 7 + \gamma 7 + \gamma^2 7 + ... = 7(\displaystyle\sum_{k=0} ^{\infty} \gamma^k) = 7 \frac{1}{1-\gamma}$

and

$G_{0} = 2 + \gamma G_{1}$

```{r}

gamma = 0.9
R2 = 7

G1 = 7 * (1/(1-gamma))  
printf(paste("G1 =", G1, sep=" "))

G0 = 2 + gamma*G1
print(paste("G0 =", G0, sep=" "))

```



